---
title: "Lizard Lengths"
author: "Deanna Elliott"
date: "11/26/2021"
output: html_document
---

```{r setup, include=FALSE, message = FALSE, warning = FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE)
options(digits = 4)

library(tidyverse)
library(here)
library(janitor)
library(kableExtra)
library(effsize)
```

### Overview

This report provides an exploratory overview of differences in female lizard length for two species--the Colorado Checkered lizard (*Cnemidophorus teassalatus*) and the Desert Grassland lizard (*Cnemidophorus uniparens*)--observed within the Jornada Basin Long Term Ecological Research (LTER) network site from 1989 - 2006 (Lightfoot 2017). Following exploratory data visualization, species lengths are analyzed by two-sample t-test using a significance level (α) of 0.05. Differences between groups are further described by Cohen’s d effect size. All analyses are in R version 4.1.1 using RStudio version 1.4.1717.

### Data Visualization & Analysis

```{r}
# read in data

liz <- read_csv(here("data", "lizards.csv"))

# create subset by spp, sex, total_length

liz_sub <- liz %>% 
    subset(select = c(spp, sex, total_length)) %>% 
  filter(spp %in% c("CNUN", "CNTE")) %>% 
  filter(sex == "F") %>% 
   mutate(spp_full = case_when(
    spp == "CNUN" ~ "Cnemidophorus uniparens",
    spp == "CNTE" ~ "Cnemidophorus tessalatus"))
```

```{r}

# plot lengths by spp

ggplot(data = liz_sub,
       aes(x = spp_full, y = total_length, fill = spp_full)) +
   scale_fill_manual(values = c("violetred", "royalblue1")) +
  geom_dotplot(binaxis = "y", stackdir = "center", binwidth = 8) +
  stat_summary(fun.y=mean, geom="point", shape=18,size=5, color="darkgoldenrod1") +
  labs(x = "Species", y = "Total Length (mm)") +
  theme_bw() +
  theme(legend.position = "none")
```

**Figure 1. Female lizard lengths by species.** Length, sex, and species were recorded for lizards in the Jornada Basin LTER. This graph compares the sizes of female lizards of two species (*Cnemidophorus tessalatus* and *Cnemidophorus uniparens*). Overall, *C. tessalatus* length is larger, with middling overlap with *C. uniparens*. The yellow diamond in each species plot shows the mean length for that species. *Data: Lightfoot (2017)*.

.

```{r}
# find summary stats for both species 

liz_sub %>% 
  group_by(spp_full) %>%
  summarize(mean = mean(total_length, na.rm = TRUE), stdev = sd(total_length, na.rm = TRUE), sample = n()) %>% 
   kbl(col.names = c("Species", "Mean Length (mm)", "Standard Devation (mm)", "Sample Size"), caption = "<b> Table 1. Summary statistics for female lizard length.</b> Mean and standard deviation in mm, with sample sizes of the females of two lizard species observed at the Jornada Basin LTER. *Data: Lightfoot (2017)*.") %>% 
  kable_classic()
```

```{r, exploration of assumptions, include = FALSE}

# subset by species

liz_u <- liz_sub %>% 
  filter(spp == "CNUN") %>% 
  drop_na(total_length) %>% 
  pull(total_length)

liz_t <- liz_sub %>% 
  filter(spp == "CNTE") %>% 
  drop_na(total_length) %>% 
  pull(total_length)

# explore normality of data with histograms and qq plots

hist(liz_t) # normal-ish
hist(liz_u) # normal-ish

ggplot(data = liz_sub, aes(sample = total_length)) +
  geom_qq() +
  facet_wrap(~ spp)

# t data looks normal enough, u data looks pretty normal
# I can use a t-test
```

```{r}

# test for difference in means

t_test <- t.test(liz_t, liz_u)

# test for effect size

cd <- cohen.d(liz_t, liz_u)
```

.

The mean length of female *C. tessalatus* at Jornada Basin LTER is `r round(t_test$estimate[1] - t_test$estimate[2], 1)` g greater than the mean length of female *C. uniparens*; this is a 49.58% difference. If these samples were taken from populations with the same mean, there is a 4e^-10^% chance of finding sample means *at least* as different as these, which is very remote. Additionally, the means of both species' lengths do differ significantly (t-test | t(`r round(t_test$parameter, 0)`) = `r round(t_test$statistic, 2)`, *p* `r ifelse(t_test$p.value < 0.001, paste("< 0.001"), paste("=", t_test$p.value))`), and the effect size of the difference is large (Cohen's d |  d = `r round(cd$estimate, 2)`).

### Citation

**Data**: Lightfoot, D. 2017. Lizard pitfall trap data (LTER-II, LTER-III) ver 36. Environmental Data Initiative. https://doi.org/10.6073/pasta/ff37b4cace16a9943575f3fd7067064e 

.
