---
title: "Lizard Lengths"
author: "Deanna Elliott"
date: "11/26/2021"
output: html_document
---

```{r setup, include=FALSE, message = FALSE, warning = FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE)
options(digits = 4)

library(tidyverse)
library(here)
library(janitor)
library(kableExtra)
```

### Overview




### Data Analysis & Visualization

```{r}
# read in data

liz <- read_csv(here("data", "lizards.csv"))

# create subset by spp, sex, total_length

liz_sub <- liz %>% 
    subset(select = c(spp, sex, total_length)) %>% 
  filter(spp %in% c("CNUN", "CNTE")) %>% 
  filter(sex == "F") %>% 
   mutate(spp_full = case_when(
    spp == "CNUN" ~ "Cnemidophorus uniparens",
    spp == "CNTE" ~ "Cnemidophorus tessalatus"))
```

```{r}

# plot lengths by spp

ggplot(data = liz_sub,
       aes(x = spp_full, y = total_length, fill = spp_full)) +
   scale_fill_manual(values = c("violetred", "royalblue1")) +
  geom_dotplot(binaxis = "y", stackdir = "center", binwidth = 8) +
  stat_summary(fun.y=mean, geom="point", shape=18,size=5, color="darkgoldenrod1") +
  labs(x = "Species", y = "Total Length (mm)") +
  theme_bw() +
  theme(legend.position = "none")
```

**Figure 1. Female lizard lengths by species.** Length, sex, and species were recorded for lizards in the Jornada Basin LTER. This graph compares the sizes of female lizards of two species (*Cnemidophorus tessalatus* and *Cnemidophorus uniparens*). Overall, *C. tessalatus* length is larger, with middling overlap with *C. uniparens*. The yellow diamond in each species plot shows the mean length for that species. *Data: Lightfoot (2017)*.

.

```{r}
# find summary stats for both species 

liz_sub %>% 
  group_by(spp_full) %>%
  summarize(mean = mean(total_length, na.rm = TRUE), stdev = sd(total_length, na.rm = TRUE), sample = n()) %>% 
   kbl(col.names = c("Species", "Mean Length (mm)", "Standard Devation (mm)", "Sample Size"), caption = "<b> Table 1. Summary statistics for female lizard length.</b> Mean and standard deviation in mm, with sample sizes of the females of two lizard species observed at the Jornada Basin LTER. *Data: Lightfoot (2017)*.") %>% 
  kable_classic()
```
```

### Citation
